---
title: "Aral Sea Shrinkage Animation (2007-2020)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Aral Sea Animation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = FALSE,  # Prevents execution during package build
  warning = FALSE,
  message = FALSE
)
```

## Introduction
This vignette demonstrates how to create an animation of the Aral Sea's shrinkage  (2007-2020) using MODIS data from Google Earth Engine.

## Setup
```{r setup-pkgs, eval=FALSE}
library(animategee)
library(sf)
library(rgee)
```

## Step 1: Initialize GEE
```{r init-gee, eval=FALSE}
gee_init()  # Authenticate and initialize (requires GEE account)
rgee::ee_check() # Verify installation
```

## Step 2: Define Region of Interest
```{r define-region, eval=FALSE}
aral_sea<- st_as_sfc(
  "POLYGON ((58 46.5, 61 46.5, 61 44, 58 44, 58 46.5))", 
  crs = 4326
)
```

## Step 3: Fetch NDWI Data (Water Index)
```{r fetch-data, eval=FALSE}
ndwi_data <- fetch_ts(
  collection = "MODIS/006/MOD09A1",  # Surface reflectance
  band = "NDWI",                     # Normalized Difference Water Index
  region = aral_sea,
  dates = c("2007-01-01", "2020-12-31"),
  scale = 500,                       # 500m resolution
  reducer = "median"                 # Reduce cloud effects
)
```

## Step 4: Create Animation
```{r animate, eval=FALSE}
animate_gif(
  df = ndwi_data,
  output_file = "aral_sea_shrinkage.gif",
  y_col = "NDWI",
  title = "Aral Sea Water Loss (2007-2020)",
  fps = 8,                           # Frames per second
  color = "blues"                    # Color scale for water
)
```

## Interpretation
- NDWI > 0.2: Water presence (blue colors)
- NDWI < 0: Land
- **Declining values**: Water loss
- The animation will show:
  - Rapid shrinkage from 2007-2010
  - Eastern basin dries completely
  - Complete separation into North/South basins by 2015
  
## Complete Script
For immediate use:
```{r complete-script, eval=FALSE}
library(animategee)
gee_init()
aral_sea <- st_as_sfc("POLYGON ((58 46.5, 61 46.5, 61 44, 58 44, 58 46.5))")
ndwi_data <- fetch_ts("MODIS/006/MOD09A1", "NDWI", aral_sea, 
                     c("2007-01-01", "2020-12-31"), scale = 500)
animate_gif(ndwi_data, "aral_sea.gif", y_col = "NDWI", 
           title = "Aral Sea Shrinkage", fps = 8, color = "blues")
```
